---
title: "ANA515_Assignment 4"
author: "Sher Mayne"
date: "4/30/2021"
output:
  pdf_document: default
---

##Business Goal - Boston is one of the cities that has a huge concentration of college students and young professionals. With its growing population, the housing market has been booming, leading to an exponential increase in the sale price of the homes in the area. The purpose of this research is to understand the Boston home prices in the 1970s based on various variables, focusing on the number of rooms, riverfront property, accessibility to highways, and distance to employment centers. My aim is the determine which different variables are most influential on the housing price. At the end of this research, we will be able to find out how each of the factors affects the home price in Boston.

##Dataset Information - The dataset I will be using is Boston Housing Dataset. The information in this dataset is collected by the U.S. Census Service which can be found on StatLib archive. You may download the file from url = http://lib.stat.cmu.edu/datasets/boston

##Cleaning and Importing Dataset - The data was cleaned manually by having all 14 variables on the same row and inputing the column names. This is to allow R to import the txt file properly.
```{r message=FALSE}
library(readr)
boston_housing<-read_delim("Boston_Housing.txt", delim = " ",skip = 21)
```
```{r}
head(boston_housing,3)
```
#Change Variables' Class Type
```{r}
boston_housing$CRIM<-as.numeric(boston_housing$CRIM)
boston_housing$ZN<-as.numeric(boston_housing$ZN)
boston_housing$INDUS<-as.numeric(boston_housing$INDUS)
boston_housing$CHAS<-as.numeric(boston_housing$CHAS)
boston_housing$NOX<-as.numeric(boston_housing$NOX)
boston_housing$RM<-as.numeric(boston_housing$RM)
boston_housing$AGE<-as.numeric(boston_housing$AGE)
boston_housing$DIS<-as.numeric(boston_housing$DIS)
boston_housing$RAD<-as.numeric(boston_housing$RAD)
boston_housing$TAX<-as.numeric(boston_housing$TAX)
boston_housing$PTRATIO<-as.numeric(boston_housing$PTRATIO)
boston_housing$B<-as.numeric(boston_housing$B)
boston_housing$LSAT<-as.numeric(boston_housing$LSAT)
boston_housing$MEDV<-as.numeric(boston_housing$MEDV)
print(sapply(boston_housing,class))
```
#Change Column Names to Lowercase
```{r message=FALSE}
library(dplyr)
```
```{r}
boston_housing<-rename_all(boston_housing,tolower)
head(boston_housing, 3)
```


##Describing Dataset
This dataset has `r ncol(boston_housing)` columns and `r nrow(boston_housing)` rows.
```{r results=FALSE}
ncol(boston_housing)
nrow(boston_housing)
```
#Data Description
```{r}
variables<-colnames(boston_housing)
details<-c("per capita crime rate by town",
           "proportion of residential land zoned for lots over 25,000 sq.ft.",
           "proportion of non-retail business acres per town",
           "Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)",
           "nitric oxides concentration (parts per 10 million)",
           "average number of rooms per dwelling",
           "proportion of owner-occupied units built prior to 1940",
           "weighted distances to five Boston employment centres",
           "index of accessibility to radial highways",
           "full-value property-tax rate per $10,000",
           "pupil-teacher ratio by town",
           "1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town",
           "% lower status of the population",
           "Median value of owner-occupied homes in $1000's")
print(data_description<-tibble(variables,details))
``` 
#Summary Statistics of Dataset
```{r}
print(sboston_housing<-(summary(boston_housing)))
```

##Data Preparation - As mentioned above, cleaning was done before importing the data as it was not possible to import the dataset without errors. I will only be using 4 variables against the home price in my analysis.
```{r}
colnames(boston_housing)
proj_bhm<-boston_housing[c(4,6,8,9,14)]
head(proj_bhm, 3)
```
#Rearranging Columns
```{r}
proj_bhm<-proj_bhm[, c(5, 2, 1, 3, 4)]
head(proj_bhm, 3)
```
##Average Home Price in Boston
```{r}
smedv<-summary(proj_bhm$medv)
srm<-summary(proj_bhm$rm)
schas<-summary(proj_bhm$chas)
sdis<-summary(proj_bhm$dis)
srad<-summary(proj_bhm$rad)
sum_stats<-c("Min","1stQu","Median","Mean","3rdQu","Max")
print(summarydata<-tibble(sum_stats,smedv,srm,schas,sdis,srad))
cchas<-c("non-Riverfront","Riverfront")
tibble(cchas,table(boston_housing$chas))
```
As we can see from the summary table, the average price of a Boston home in the 1970s is `r mean(proj_bhm$medv)`, where it has `r mean(proj_bhm$rm)` rooms, not by Charles River, `r mean(boston_housing$dis)` weighted distance to five Boston employment centres, and `r mean(boston_housing$rad)` of index accessibility to radial highways.
```{r results=FALSE}
mean(boston_housing$rm)
mean(boston_housing$dis)
mean(boston_housing$rad)
mean(boston_housing$medv)
```
We can look deeper into the result above by plotting them out.
```{r}
bhmmedv<-proj_bhm$medv
bhmrm<-proj_bhm$rm
bhmchas<-proj_bhm$chas
bhmdis<-proj_bhm$dis
bhmrad<-proj_bhm$rad
hist(bhmmedv,main="Home Prices in Boston",xlab="Home Price in $1000's",xlim=c(0,50),col=("lightblue"),freq=FALSE)
plot(bhmrm,bhmmedv,main="Home Price vs Number of Rooms",xlab="Number of Rooms",ylab="Home Price in $1000's",pch=20,col=("darkblue"))
plot(bhmdis,bhmmedv,main="Home Price vs Weighted Distance to 5 Employment Centres",xlab="Weighted Distance",ylab="Home Price in $1000's",pch=20,col=("darkblue"))
plot(bhmrad,bhmmedv,main="Home Price vs Index Accessibility to Radial Highways",xlab="Index Accesibility",ylab="Home Price in $1000's",pch=20,col=("darkblue"))
barplot(table(proj_bhm$chas),main="Home Price vs Charles Riverfront Property",xlab="Charles River (0 = no, 1 = yes)",ylab="Home Price in $1000's",col=("darkblue"))
```

##Correlation Matrix
```{r}
print(round(cor(proj_bhm),4))
```
Based on the correlation matrix table, the variable which has the strongest association with the home price is the number of rooms with a value of 0.6954, while the variable which has the weakest association with the home price is Charles Riverfront property with a value of 0.1753. There is no evidence of multicollinearity as the correlation among the four variables of rm, chas, dis, and rad  are not larger than 0.7

##Multiple Regression Model
```{r}
library(olsrr)
ols_regress(bhmmedv ~ bhmrm + bhmchas + bhmdis + bhmrad, data = proj_bhm)
```
I formulated the estimated regression equation using Ordinary Least Squares function. Price = -27.455 + 8.264bhmrm + 4.239bhmchas + 0.053bhmdis - 0.256bhmrad

Let us interpret the estimated regression equation. For an increase of 1 bedroom, we expect the home price to increase by $8264. When a home is a Charles Riverfront property, the home price is $4239 higher relative to a non-waterfront property. For a 1 unit increase weighted distances to five Boston employment centres, the home price increases by $5.3. For a one unit increase in index of accessibility to radial highways, the home price decreases by $256. The slope coefficient for number of rooms, Charles Riverfront property, and index of accessibility to radial highways are statistically significant as p-value < 0.05. 